\textls[-0]{\thetitle}

### 2023-03-15 related-work.tex
incorporate some of Zhixuan's comment
took a pass over related work sec

%DTC is as a source of inspiration for may approaches
%\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}\EDJreply{My understanding is,
 it is really an intrinsic typed version of MTC. It uses different set of tools (like containers (a la altenkirch, can be considered as an alternative of W-type) and stuff) but it is still a design pattern paper.
 But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach).}
%\cite{vdr2022intrinsically}

All of these approaches are largely extra\-linguistic, in that they work
within the confine of the language offered by a proof assistant, which
comes with trade-offs.
On the one hand,
they can be conveniently distributed as libraries, and the encoding can
be more easily adapted for new purposes.
For example, MTC~\cite{delaware2013} has been applied to implementing
composable program adverbs~\cite{liwei2022} and been adapted to support
unanticipated feature extensions that require type changes
\cite{delaware2013b}.

@@

\citet{forsta2020} introduce \emph{Coq à la carte}.
It still follows DTC, but rather than embracing DTC's use of generic
fixed points, it considers specific instantiations instead.
The resulting mechanism appears more streamlined than prior \emph{à la
carte} approaches particularly for its extensive tool support for generating boilerplate code.
%
But even with the tool support, components (e.g., "substlem") of
individually developed feature extensions have to be composed separately
by invoking the tool.
%
%The approach does not seem to fully address the expression problem,
%as \citet{liwei2022} observe, in that further extending a composed
%mechanization causes (a small amount of) boilerplate code that glues definitions
%to be regenerated.

@@

\noindentparagraph{Solutions to the expression problem} \!abound.
% \EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
Almost all involve some form of either explicit or implicit parameterization
as extensibility hooks.
Our approach is the first that applies family polymorphism~\cite{ernst2001family},
an idea from OO research,
to the context of mechanized proofs.
%Families in \Lang are second-class, in that they are not considered
%terms by themselves.
%Families as modeled in \TT (namely linkages) are first-class and thus more flexible.\EDJ{A more unified way is to consider Coq's Family (our surface syntax) second class corresponds to Inheritance Judgement/Linkage Transformer (secondclass in type theory). Linkage transformer is the helper function to construct first class linkage easily. Even non-derived (surface syntax) family can be considered in this way because they are just inheritance on empty family.}
%The distinction corresponds to class-based vs.\ object-base family
%polymorphism as summarized before~\cite{cdnw07-tribe}.

>>>>

%DTC is as a source of inspiration for may approaches
%\cite{delaware2013,delaware2013b,schwaab2013modular,keuchel2013generic,forsta2020}.
The original DTC encoding requires type-level general recursion that
fails the strict positivity check imposed by proof assistants including Coq and Agda.
\citet{delaware2013} introduce \emph{meta-theory à la carte} (MTC):
it overcomes the problem by using Church encodings for
data types and using Mendler-style folds for evaluation,
though it requires \lsti{Set} impredicativity.
Feature composition is automated through heavy use of type classes.
The framework is implemented as a Coq library.
\citet{schwaab2013modular} adapt DTC to Agda by considering a restricted
class of functors that admit least fixed points.
\citet{keuchel2013generic} use datatype-generic programming techniques
for the underlying representation of type-level fixed points and
avoid \lsti{Set} impredicativity.
%\Citet{vdr2022intrinsically} mechanize intrinsically-typed
%target language based on container and signature \cite{altenkirch2015container}. 
%Thanks to the nature of intrinsically-typed abstract syntax trees (ASTs) and Agda's advanced unification mechanism, the authors were able to eliminate a significant amount of code that dealt with "impossible" cases.
%\YZ{What can we say about Intrinsically-Typed Definitional Interpreters a La Carte?}\EDJreply{My understanding is,
% it is really an intrinsic typed version of MTC. It uses different set of tools (like containers (a la altenkirch, can be considered as an alternative of W-type) and stuff) but it is still a design pattern paper.
% But good news is, container is closer to Wtype(inductive type) than the endofunctor encoding (classic MTC approach)}\EDJreply{I also add one sentence after Coq a la carte. You can also check... I will add bib later. Interestingly, Intrinsic type doesn't refer Coq a la carte.}

All of these approaches are largely extra\-linguistic, in that they work
within the confine of the language offered by a proof assistant, which
comes with trade-offs.
On the one hand,
they can be conveniently distributed as libraries, and the encoding can
be more easily adapted for new purposes.
For example, MTC~\cite{delaware2013} has been applied to implementing
composable program adverbs~\cite{liwei2022} and been adapted to allow
unanticipated feature extensions that require type changes
\cite{delaware2013b,vdr2022intrinsically}.

@@

\citet{forsta2020} introduce \emph{Coq à la carte}.
It still follows DTC, but rather than embracing DTC's use of generic
fixed points, it considers specific instantiations instead.
The resulting mechanism appears more streamlined than prior \emph{à la
carte} approaches particularly for its extensive tool support for generating boilerplate code.
%
But even with the tool support, components (e.g., "substlem") of
individually developed feature extensions have to be composed separately
by invoking the tool.
%The approach does not seem to fully address the expression problem,
%as \citet{liwei2022} observe, in that further extending a composed
%mechanization causes (a small amount of) boilerplate code that glues definitions
%to be regenerated.

@@

\noindentparagraph{The expression problem.} Solutions abound.
% \EDJ{I have been meaning to ask but is this ``abound'' you intentionally put outside of the boldfont?}\YZreply{Yes}
Almost all involve some form of either explicit or implicit parameterization
as extensibility hooks.
Our approach is the first that applies family polymorphism~\cite{ernst2001family},
mostly seen in OO languages,
to the context of mechanized proofs.
%Families in \Lang are second-class, in that they are not considered
%terms by themselves.
%Families as modeled in \TT (namely linkages) are first-class and thus more flexible.\EDJ{A more unified way is to consider Coq's Family (our surface syntax) second class corresponds to Inheritance Judgement/Linkage Transformer (secondclass in type theory). Linkage transformer is the helper function to construct first class linkage easily. Even non-derived (surface syntax) family can be considered in this way because they are just inheritance on empty family.}
%The distinction corresponds to class-based vs.\ object-base family
%polymorphism as summarized before~\cite{cdnw07-tribe}.
